---
title: "ec1207"
author: "Revanth"
date: "2024-03-24"
output: html_document
---

```{r}
library(psych)
library(nFactors)
library(datasets)
library(readr)
```

```{r}
# Load Iris dataset
data(iris)
```

```{r}
# Exploring dataset
str(iris)
summary(iris)
```

```{r}
# Performing PCA to understand component analysis
iris_pca <- prcomp(iris[, -5], scale. = TRUE)

# Summary of the PCA
summary(iris_pca)

```

```{r}
# Biplot of the PCA
biplot(iris_pca)
```

# Q1. Decide how many Factors are ideal for your dataset

```{r}
# Performing parallel analysis to decide the number of factors
iris_numeric <- iris[, 1:4]
iris_parallel <- fa.parallel(iris_numeric, fa = "both", n.iter = 1000, n.obs = nrow(iris_numeric))

```

```{r}
# Plot parallel analysis results
plot(iris_parallel)
```

```{r}
# Perform factor analysis with 2 factors
fit.fa <- fa(iris_numeric, nfactors = 2, rotate = "varimax")

# Perform factor analysis with 3 factors
fit.fa3 <- fa(iris_numeric, nfactors = 3, rotate = "varimax")
```

# Q2. Explain the output for your factor model

```{r}
# Explain the factor model
summary(fit.fa)

summary(fit.fa3)
```

## The output provides information about the factor analysis model:
## The standardized loadings (pattern matrix) show the correlation between each variable and the extracted factors. Higher loadings indicate a stronger relationship.
## The "SS loadings" row shows the sum of squared loadings for each factor.
## The "Proportion Var" row shows the proportion of variance explained by each factor.
## The "Cumulative Var" row shows the cumulative proportion of variance explained by the factors.
## The chi-square statistic tests the hypothesis that the specified number of factors is sufficient to explain the correlations among the variables.

# Q3. Show the columns that go into each factor 

```{r}
# Show the columns that go into each factor
loadings(fit.fa)
loadings(fit.fa3)
```

## From the pattern matrix in the output, we can see that:

## Factor 1 is strongly associated with the "Sepal.Length" and "Sepal.Width" columns.
## Factor 2 is strongly associated with the "Petal.Length" and "Petal.Width" columns.

```{r}
# Extract factor loadings
factor_loadings <- fit.fa$loadings

factor_loadings3 <- fit.fa3$loadings

```

# Q4. Perform some visualizations using the factors

```{r}
# Plot factor loadings
plot(factor_loadings, type = "n", xlim = c(-1, 1), ylim = c(-1, 1), xlab = "Factor 1", ylab = "Factor 2")
text(factor_loadings[, 1], factor_loadings[, 2], labels = rownames(factor_loadings), cex = 0.8)
abline(h = 0, v = 0, col = "red", lty = 2)


```



```{r}
# Create biplot
biplot(fit.fa, choices = c(1, 2), labels = rownames(iris_numeric))
```

## This code extracts the factor scores from the factor analysis output and plots them on a scatterplot. The x-axis represents Factor 1 (associated with sepal dimensions), and the y-axis represents Factor 2 (associated with petal dimensions)

```{r}
# Component analysis
fa.plot(fit.fa3)
fa.diagram(fit.fa3)
```
```{r}
vss(iris[,-5])
```
```{r}

```
```{r}

```