---
title: "Social_media"
author: "Revanth"
date: "2024-04-14"
output: html_document
---

```{r}
media <- read.csv("c:/users/admin/desktop/media.csv", row.names=1)

str(media)

colnames(media) <- c("Instagram", "LinkedIn", "Snapchat", "Twitter", "Whatsapp", "Youtube", "OTT", "Reddit", "Trouble_sleep", "Mood", "Tired_morning")

```

# 1. Model Development

```{r}
# model

# Performing multiple regression on media dataset
fit <- lm(Trouble_sleep~Instagram + Snapchat + Twitter, data=media) # LM- ASSUMES IT TO BE IT LINEAR, y(mpg)- many variables-output-residuals means error -median of the errors- median means real minus predicted

```

```{r}
#show the results
summary(fit)

```

# 2. Model fit 

### Insights:

#### The statistical significance of each coefficient is assessed through their corresponding t-values and p-values. A low p-value (< 0.05) suggests that the predictor variable is statistically significant in explaining the variability in trouble with sleep.

```{r}
fit <- lm(Trouble_sleep~Instagram + Snapchat + Twitter, data=media)
#removed the values that were 0.05

# Useful Helper Functions
coefficients(fit)

```

```{r}
library(GGally)
ggpairs(data=media, title="Social media Data")

```

```{r}
confint(fit,level=0.95) #confidence interval
fitted(fit)
```

# 3. Residual Analysis

```{r}
residuals(fit)
#Anova Table
anova(fit)
vcov(fit)
cov2cor(vcov(fit))
temp <- influence.measures(fit)#get to know abt outliers of data - makes * at the end
temp
plot(fit)
```

```{r}
library(car)
outlierTest(fit)#current one shows outlier 
leveragePlots(fit) # leverage plots
# Influential Observations
# added variable plots
avPlots(fit)
```

```{r}
# Cook's D plot
# identify D values > 4/(n-k-1)
cutoff <- 4/((nrow(media)-length(fit$coefficients)-2))
plot(fit, which=4, cook.levels=cutoff)
```

```{r}
# Influence Plot
influencePlot(fit, id.method="identify", main="Influence Plot", sub="Circle size is proportial to Cook's Distance" )
```

```{r}
# Normality of Residuals
# qq plot for studentized resid
qqPlot(fit, main="QQ Plot")

```

### Insights : Residuals represent the differences between the observed values of the response variable (Trouble_sleep) and the values predicted by the regression model.

#### For example, observations like masinl, Patty, tl868, trave, ds2134, Harvey, ki567, MVA37@S have negative residuals, suggesting that the predicted trouble with sleep is higher than the actual trouble with sleep for these individuals. Conversely, observations like peace, 19!@s, ak2001, vp1234, 15801 have positive residuals, indicating that the predicted trouble with sleep is lower than the actual trouble with sleep for these individuals.

```{r}
library(MASS)
sresid <- studres(fit)
hist(sresid, freq=FALSE, main="Distribution of Studentized Residuals")
xfit<-seq(min(sresid),max(sresid),length=40)
yfit<-dnorm(xfit)
lines(xfit, yfit)
#Non-constant Error Variance
# Evaluate homoscedasticity

```

```{r}
# non-constant error variance test
ncvTest(fit)

```

```{r}
# plot studentized residuals vs. fitted values
spreadLevelPlot(fit)

```

```{r}
#Multi-collinearity
# Evaluate Collinearity
vif(fit) # variance inflation factors
sqrt(vif(fit)) > 2 

```

```{r}
#Nonlinearity
# component + residual plot
crPlots(fit)

```

```{r}
# Ceres plots
ceresPlots(fit)
```

```{r}
# Test for Autocorrelated Errors
durbinWatsonTest(fit)
```
# Model Accuracy

```{r}
library(gvlma)
gvmodel <- gvlma(fit)
summary(gvmodel)
fit
summary(fit)
fit1 <- fit
fit2 <- lm(Trouble_sleep~Instagram + Snapchat + Twitter, data = media)
```


```{r}
# compare models
anova(fit1, fit2)
step <- stepAIC(fit, direction="both")
step$anova # display results
library(leaps)
leaps<-regsubsets(Trouble_sleep~Instagram + Snapchat + Twitter,data=media,nbest=10)

```


```{r}
# Displaying results
plot(leaps)
plot(leaps,scale="r2")
plot(leaps,scale="bic")
summary(leaps)
```

### Insights: The comparison between Model 1 and Model 2 suggests that including or excluding the variables Instagram, Snapchat, and Twitter does not significantly impact the model's ability to explain the variability in Trouble_sleep.

# Model Prediction

```{r}
library(relaimpo)
calc.relimp(fit,type=c("lmg","last","first","pratt"),
            rela=TRUE)
# Bootstrap Measures of Relative Importance (1000 samples)
boot <- boot.relimp(fit, b = 1000, type = c("lmg",
                                            "last", "first", "pratt"), rank = TRUE,
                    diff = TRUE, rela = TRUE)
booteval.relimp(boot) # print result
plot(booteval.relimp(boot,sort=TRUE)) # plot result
#https://rpubs.com/davoodastaraky/mtRegression
summary(fit)
predict.lm(fit, data.frame(Instagram =0.2 ,Snapchat=0.1,Twitter=0.5) )

```

### Insights:

#### The predicted value of Trouble_sleep is approximately 0.2391868.

#### Interpretation: This means that when the values of Instagram is 0.2, Snapchat is 0.1, and Twitter is 0.5, the model predicts that the value of Trouble_sleep will be approximately 0.2391868.